---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Teddy Pierson"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
cookies <- read.csv("choc_chip_cookie_ingredients.csv")
library(reticulate)
py_require(c("pandas"))
```

```{python}
import pandas as pd
cookies = pd.read_csv("choc_chip_cookie_ingredients.csv")
```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### GeneRating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. Which ingredients have the highest Rating?
2. Does a higher quantity of ingredient lead to a higher Rating?
3. Which measurement (cups, teaspsoon, tablespoon) occurs most often?
4. Which recipe source has the highest average Rating?
5. How many Ratings have a value of NA?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library(skimr)
skim1 <- skim(cookies)
skim2 <- skim_tee(cookies)
skim3 <- skim_without_charts(cookies)
library(reticulate)
py_require(c("skimpy", "numpy"))
```

```{python}
from skimpy import skim
```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

I found the range of word lengths in the data, which was interesting. I also saw the histogram for the data, and I displayed the data without the histogram by using skim_without_charts

### GeneRating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients
.
```{r}
library(dplyr)
total_recipes <- n_distinct(cookies$Recipe_Index)
ingredient_props <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%     
  count(Ingredient, sort = TRUE) %>%    
  mutate(proportion = n / total_recipes) %>%
  slice_head(n = 20)

ingredient_props
```

```{python}
import pandas as pd
 
total_recipes = cookies["Recipe_Index"].nunique()
 
ingredient_props = (
    cookies[["Recipe_Index", "Ingredient"]]
    .drop_duplicates()
    .groupby("Ingredient")
    .size()
    .reset_index(name="count")
)
 
ingredient_props["proportion"] = ingredient_props["count"] / total_recipes
 
ingredient_props = ingredient_props.sort_values(
    by="count", ascending=False
).head(20)
 
ingredient_props
```


[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.

```{r}
library(dplyr)

ingredient_counts <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%  
  count(Ingredient)

rare_ingredients <- ingredient_counts %>%
  filter(n < 20) %>%
  pull(Ingredient)

paste(rare_ingredients, collapse = ", ")
```

```{python}
ingredient_counts = (
    cookies[["Recipe_Index", "Ingredient"]]   
    .drop_duplicates()
    .groupby("Ingredient")
    .size()
    .reset_index(name="count")
)
 
rare_ingredients = ingredient_counts[
    ingredient_counts["count"] < 20
]["Ingredient"]
 
print(", ".join(rare_ingredients))

```


### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.

```{r}
library(tidyverse)
library(ggplot2)
library(scales)

cookies_clean <- cookies %>%
  filter(!is.na(Rating))

total_recipes_clean <- n_distinct(cookies_clean$recipe)

cookies_clean %>%
  group_by(Ingredient) %>%
  filter(n() >= 5) %>%
  ungroup() %>%
  ggplot(aes(x = fct_reorder(Ingredient, Rating, median, na.rm = TRUE), 
             y = Rating)) +
  geom_boxplot(fill = "blue", alpha = 0.7, outlier.color = "red", outlier.alpha = 0.3) +
  coord_flip() +
  labs(
    title = "Recipe Ratings by Ingredient",
    subtitle = paste0("All Ingredients appearing in 5+ recipes"),
    x = "Ingredient",
    y = "Rating",
    caption = "Box shows median and IQR; dots are outliers"
  ) +
  theme_minimal()
```



[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

From this plot, you can see which ingredients have the best overall rating. I had to remove the NA's for the code to work, which may unfairly affect the ingredients that have fewer NA's. I also could limit the scope to ingredients that appear in 15 or 20+ recipes instead of 5. That might lead to a better result. But for now, I can conclude that pudding mix and white chocolate chip are the highest-rated ingredients
